---
title: "CF_rnaseq"
author: "Manpreet S. Katari"
date: "1/18/2021"
output: html_document
---

## 4.2 - The data
The following command downloads the dataset directly from the NCBI's GEO site. In case it doesn't work, the same Excel file is provided in this repository, under CaseStudies/RNAseq. 


```{r}
library(readxl)
data_df <- read_excel("RNASeqData.xlsx")
head(data_df)
```
The column headers have some unnecessary information so we will remove them to make it cleaner. 

```{r}
column_names = gsub("w_[0-9]*_O","w_O", colnames(data_df), perl=T)
column_names = gsub("m_[0-9]*_O","m_O", column_names, perl=T)
column_names = gsub("w_[0-9]*_H","w_H", column_names, perl=T)
column_names = gsub("m_[0-9]*_H","m_H", column_names, perl=T)

colnames(data_df) = column_names
```


### 4.3.1 normalization

The first 5 columns are gene descriptions and identifiers. In this exercise we will start with the **raw** data. First we obtain all the columns that contain the word **raw** and use them to perform our normalization.

As discussed in the book, there are many different methods of normalizing the data and we are choosing to use a very simple. For each gene g, this approach first maps the raw count of g to the relative abundance of g compared to all other genes for this patient; second, it multiplies that relative abundance by one million.

```{r}
raw_cols = grep("Raw", colnames(data_df), value = T)

raw_cols_sums = colSums(data_df[,raw_cols])

norm_cpm = t(t(data_df[,raw_cols]) * (1000000/raw_cols_sums))

head(norm_cpm[,1:5])
```

```{r}

norm_cpm_df = as.data.frame(norm_cpm)
dim(norm_cpm_df)

norm_cpm_df = cbind(norm_cpm_df, data_df[,1:9])
head(norm_cpm_df)
```
```{r}
colSums(norm_cpm_df[,1:5])

```

### 4.3.2 Remove genes that contain zeroes

```{r}
gene_with_min_greaterthan_zero = which(apply(norm_cpm_df[,raw_cols], 1, min) > 0)


```

Keep only genes with no zeroes

```{r}
data_df_subset = norm_cpm_df[gene_with_min_greaterthan_zero,]
dim(data_df_subset)

```

```{r}
number_of_zeroes_confirm = data_df[,raw_cols]==0
sum(number_of_zeroes_confirm)
number_of_zeroes_confirm = data_df_subset[,raw_cols]==0
sum(number_of_zeroes_confirm)

```

## 4.4 Distinguishing sick patients from healthy

```{r}
healthy = grep("HC", colnames(data_df_subset), value=T)
cf_base = grep("Base", colnames(data_df_subset), value=T)
cf_v2 = grep("V2", colnames(data_df_subset), value=T)
healthy
```

The approach to determining genes that are differentially expressed is similar to the one we used for determining diets that are different by comparing the difference of the final weights. However, we are not looking for an increase in the mean difference. Rather we are looking for a change and we are not sure if the gene will increase or decrease due to  sickness or health.

For each gene g, we take the mRNA expression values of one group (e.g. healthy patients) and compare it with the expression values of gene g among another group (e.g. untreated cystic fibrosis patients).

To evaluate significance (i.e. calculate p-values), we check how often a shuffled set of values (shuffling  ignores whether a patient is in one group or another) shows a difference that is greater in absolute value than the observed difference. If that happens rarely, then the p-value is low.


```{r}

shuffle <- function(grps) {
  	# grps is a list of elements where each element contains values from different groups

  num_grps = length(grps)
  pool=numeric()
  
  for ( i in 1:num_grps) {
    pool = c(pool, grps[[i]])
  }
  
  pool= sample(pool, length(pool), replace=F)
  
  new_grps=list()
  start_index=1
  end_index=1
  for (i in 1:num_grps) {
    end_index = start_index + length(grps[[i]]) - 1
    new_grps[[i]]=pool[start_index:end_index]
    start_index = end_index+1
  }
  return(new_grps)
}



```


```{r}
logFC <- function(grpA, grpB){
    return(log2( (sum(grpB)/length(grpB)) / (sum(grpA) / length(grpA)))) 
}


```

```{r}

lfc2meansig = function(grpA, grpB) {
  observed_mean_diff = logFC(grpA,grpB)

  count=0
  num_shuffles=100000

  samples = list(grpA=grpA, grpB=grpB)
  
  for (i in 1:num_shuffles) {
    new_samples = shuffle(samples)
    mean_diff = logFC(new_samples[[1]], new_samples[[2]])
    if (abs(observed_mean_diff) <= abs(mean_diff)) {
      count = count+1
    }
  }
  
  return(count/num_shuffles)

}



```

```{r}

avg_diff_sig_test = function( df_norm_hc, df_norm_cf  ) {
  all_p_sig_values=numeric()
  for (i in 1:nrow(df_norm_cf)) {
    all_p_sig_values[i] = lfc2meansig(as.numeric(df_norm_hc[i,]),
                                   as.numeric(df_norm_cf[i,]))
  }
  return(all_p_sig_values)
}


```

```{r eval=F}
set.seed(123)
norm_p_values = avg_diff_sig_test(data_df_subset[healthy],
                                  data_df_subset[cf_base])
head(norm_p_values)
saveRDS(norm_p_values, "norm_p_values_unpaired_raw.rds")
```

```{r echo=F}
norm_p_values = readRDS("norm_p_values_unpaired_raw.rds")
head(norm_p_values)
```

```{r}
sig_genes = which(norm_p_values < 0.05)

print(paste("Total number of genes that pass p-value threshold of 0.05:", length(sig_genes), sep=" "))


```
```{r}
sig_gene_names = data_df_subset$ID[sig_genes]
head(sig_gene_names)

```

### 4.4.1 Bonferroni correction

In this method, we divide the threshold by the number of genes. So, if we take a threshold of 20% (or 0.2), the p-values that would pass this threshold would be `0.2/num_of_genes`. If any gene g had a p-value less than or equal to 1 in 100,000, then that would satisfy the 0.2 Bonferonni threshold because 0.2/15250 <= 2/100000

```{r}
bon_sig_genes = which(norm_p_values < (0.2/length(norm_p_values)))
print(paste("Number of genes that pass bonferroni correction threshold of 0.2: ", length(bon_sig_genes), sep=" "))
```

```{r}
bon_sig_genes_names = data_df_subset$ID[bon_sig_genes]
head(bon_sig_genes_names)

```

### 4.4.2 Benjamini–Hochberg procedure


The Benjamini-Hochberg method of creating a set of genes that have a given false discovery rate starts by listing the individual gene p-values in ascending order, from smallest to largest. The smallest p-value has a rank  $i = 1$, the next smallest has $i = 2$, etc. Compare each individual p-value to its Benjamini-Hochberg critical value, $(i/m)Q$, where $i$ is the rank, $m$ is the total number of tests (15,250, as above), and $Q$ is the false discovery rate the user can tolerate, say 5%.  The gene associated with the largest p-value $P$ that has $P<(i/m)Q$ should be included as should all of the genes associated with  p-values less than $P$ (including that aren't less than their Benjamini-Hochberg critical value). 6,082 of the genes having the lowest p-values passed this new cutoff.

```{r}
benjamin_hochberg_corrected_p_values<-function(p_values, fdr){
  num_genes = length(p_values)
  sorted_indices = order(p_values, decreasing = F)
  sorted_p_values = sort(p_values, decreasing = F)
  
  benjamin_hochberg_critical_values = numeric()
  for (i in num_genes:1) {
    benjamin_hochberg_critical_values[i]= (i/num_genes)*fdr
    if (sorted_p_values[i] < benjamin_hochberg_critical_values[i]) {
      return(list(i=i, sorted_indices = sorted_indices))
    }
  }
}

i_sorted_indices = benjamin_hochberg_corrected_p_values(norm_p_values, 0.05)

print(i_sorted_indices$i)
```
## An alternative way to adjust with Benjamini Hochberg method.

```{r}
bh_sort = p.adjust(norm_p_values, "BH")

fdr_sig_genes_names = data_df_subset$ID[i_sorted_indices$sorted_indices[1:i_sorted_indices$i]]
```

## Violin Plots

The data we loaded is in a **wide** format where the values for each gene is presented in a different row.
We will **melt** the data to create a **long** format where each value is in a separate row and every row is a specific combination of gene and sample. This format will prove useful in creating violin plots using the seaborn package.

```{r}
library(reshape2)
norm_hc_cf_with_genename = c("ID", healthy)
norm_hc_cf_with_genename_melt = melt(data_df_subset[,norm_hc_cf_with_genename], id="ID")
norm_hc_cf_with_genename_melt$variable = "HC"

norm_cf_with_genename = c("ID", cf_base)
norm_cf_with_genename_melt = melt(data_df_subset[,norm_cf_with_genename], id="ID")
norm_cf_with_genename_melt$variable = "CF"

norm_hc_cf_with_genename_melt = rbind(norm_hc_cf_with_genename_melt,norm_cf_with_genename_melt)

head(norm_hc_cf_with_genename_melt)
```


```{r}
library(ggplot2)
library(dplyr)
genestoplot = filter(norm_hc_cf_with_genename_melt, ID %in% bon_sig_genes_names[4])

ggplot(genestoplot) +
  geom_violin(mapping=aes(x=variable, y=value, fill=variable)) +
  geom_boxplot(mapping=aes(x=variable,y=value), width=0.1, alpha=0.2)

```

## 4.5 Healthy vs CF confidence interval

```{r}

# We will use the genes that passed bonferroni correction cutoff.
data_df_subset_bon = filter(data_df_subset, ID %in% bon_sig_genes_names)
df_norm_hc_bon_sig = data_df_subset_bon[,healthy]
df_norm_cf_bon_sig = data_df_subset_bon[,cf_base]
print(dim(df_norm_cf_bon_sig))
print(dim(df_norm_hc_bon_sig))


```


```{r}
bootstrap = function(x) {
  samp_x = sample(x, length(x), replace=T)
}


```

```{r}

diff2meanconf = function(grpA, grpB, num_resamples) {
  samples = list(grpA, grpB) 
    a = 1
    b = 2
    

    out = numeric()                # will store results of each time we resample
    for (i in 1:num_resamples){
        # get bootstrap samples for each of our groups
        # then compute our statistic of interest
        # append statistic to out
        bootstrap_samples = list()  # list of lists
        for (j in 1:length(samples)){
            bootstrap_samples[[j]]=bootstrap(samples[[j]])
        }

        # now we have a list of bootstrap samples, run meandiff
        out[i]=logFC(bootstrap_samples[[a]], bootstrap_samples[[b]])
    }    
    
    return(out)
    
}


```


```{r}
diff2meanJustconf<-function(out, conf_interval){
    tails = (1 - conf_interval) / 2
    num_resamples = 1000
    lower_bound = ceiling(num_resamples * tails)
    upper_bound = floor(num_resamples * (1 - tails))
    return(c(out[lower_bound], out[upper_bound]))
  
}


```

```{r}
set.seed(123)

all_mean_diff = list()

for ( i in 1:nrow(df_norm_hc_bon_sig)) {
  all_mean_diff[[i]]=diff2meanconf(df_norm_hc_bon_sig[i,],
                                   df_norm_cf_bon_sig[i,],
                                   1000)
}


```

```{r}
low_high = list()

for (i in 1:length(all_mean_diff)) {
  low_high[[i]] = diff2meanJustconf(all_mean_diff[[i]], 0.9)
}


```

```{r}
abs_conf = lapply(low_high, abs)
max_conf = lapply(abs_conf, max)
```

```{r}
hist(unlist(low_high), breaks=30)
hist(unlist(abs_conf))
hist(unlist(max_conf))

```

```{r}
top10_ci_index = order(unlist(max_conf), decreasing = T)[1:10]

top10_ci_genes = bon_sig_genes_names[top10_ci_index]
top10_ci_genes
```

```{r}

top10_ci_gene_description = data_df_subset_bon[ top10_ci_index,"Description"]
top10_ci_gene_description
```

```{r}


library(ggplot2)
library(dplyr)
genestoplot = filter(norm_hc_cf_with_genename_melt, ID %in% c("MMP9","SOCS3","ANXA3"))
dodge = position_dodge(width=.4)
ggplot(genestoplot, mapping=aes(x=ID, y=value, fill=variable)) +  
  geom_violin(position=dodge) 

```
## Figure 4.1 - Confidence interval of genes identified in published work.

```{r}

names(all_mean_diff) = bon_sig_genes_names
all_mean_diff_df = t(as.data.frame(all_mean_diff))
all_mean_diff_df = as.data.frame(all_mean_diff_df)
all_mean_diff_df$ID = rownames(all_mean_diff_df)
all_mean_diff_melt = melt(all_mean_diff_df, id.vars="ID")
head(all_mean_diff_melt)

genestoplot=all_mean_diff_melt[all_mean_diff_melt$ID %in% c("MMP9","SOCS3","ANXA3"),]


ggplot(genestoplot) + 
  geom_violin(mapping = aes(x=ID, y=value, fill=ID)) +  
  ylim(0,3)
```

## 4.6.1 Random Forest Inference

First we need to make subset of the original dataframe to contain only our top 10 genes. We will than separate them into the healthy and cf dataframes so we can transpose them and add the corresponding labels. We transpose the data because we will be using the gene expression values as the variables to predict which sample/patient is healthy or has cystic fibrosis. 

```{r}

data_df_subset_bon_top10ci = subset(data_df_subset_bon, ID %in% top10_ci_genes) 


df_norm_hc_bon_sig_top10ci = data_df_subset_bon_top10ci[healthy]
df_norm_cf_bon_sig_top10ci = data_df_subset_bon_top10ci[cf_base]
rownames(df_norm_hc_bon_sig_top10ci) = data_df_subset_bon_top10ci$ID
rownames(df_norm_cf_bon_sig_top10ci) = data_df_subset_bon_top10ci$ID

print(dim(df_norm_hc_bon_sig_top10ci))
print(dim(df_norm_cf_bon_sig_top10ci))


```
```{r}
df_norm_healthy_T = t(df_norm_hc_bon_sig_top10ci)
df_norm_base_T = t(df_norm_cf_bon_sig_top10ci)
X = rbind(df_norm_healthy_T, df_norm_base_T)
dim(X)


```

```{r}

Y = rep(c(0,1), each=20)
table(Y)

```

```{r}
library(party)
library(caret)
XY = cbind(Y,X)

rf_results = data.frame(f1=numeric(),precision=numeric(),recall=numeric())
impvar_df = data.frame(Overall=numeric(),Gene=character())
for (i in 1:100) {
    clf = train(as.factor(Y) ~ ., 
                data=XY,
                method="cforest",
                trControl=trainControl(method="cv", number=5, 
                                       savePredictions = "final"),
                
                )
    TP = sum(clf$pred$pred == 1 & clf$pred$obs == 1)
    TN = sum(clf$pred$pred == 0 & clf$pred$obs == 0)
    FP = sum(clf$pred$pred == 1 & clf$pred$obs == 0)
    FN = sum(clf$pred$pred == 0 & clf$pred$obs == 1)
    
    rf_results[i,"precision"] = (TP / (TP + FP))
    rf_results[i,"recall"] = (TP / (TP + FN))
    rf_results[i,"f1"] = (2 * rf_results[i,"precision"] * rf_results[i,"recall"]) / (rf_results[i,"precision"] + rf_results[i,"recall"]  )

    varimptemp = varImp(clf)$importance
    varimptemp$Gene = rownames(varimptemp)
    impvar_df = rbind(impvar_df, varimptemp)
    
}    
    

```

```{r}
rf_results_melt = melt(rf_results)

ggplot(rf_results_melt) +
  geom_violin(mapping = aes(x=variable,y=value,fill=variable)) +
  ylim(0,1)

```

```{r}
ggplot(impvar_df) +
  geom_violin(mapping = aes(x=Gene, y=Overall, fill=Gene ), scale="width") +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

